---
title: "Geography 176A"
author: "[Stone Shi](https://mycraysh.github.io/StoneGIS/)"
subtitle: 'Lab 03: Distances and the Border Zone'
output:
  html_document:
    theme: journal
---
```{r library,include=FALSE}
# SPDS
library(tidyverse)
library(sf)
library(units)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)

```

#Question 1

#1.1 Define a Projection
```{r,message=FALSE, warning=FALSE}
#This projection will use the North America Equidistant Conic.

eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```

#1.2 Get USA state boundaries
```{r,message=FALSE, warning=FALSE}
mapdata = USAboundaries::us_states(resolution = "low") %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) %>% 
  st_transform(eqdc)
```

#1.3 - Get country boundaries for Mexico, the United States of America, and Canada
```{r,message=FALSE, warning=FALSE}
countries = rnaturalearth::countries110
countries_sf = st_as_sf(countries) %>%
  filter(admin %in% c("Mexico","United States of America", "Canada")) %>%
  st_transform(eqdc)
```

#1.4 - Get city locations from the CSV file
```{r,message=FALSE, warning=FALSE}
uscities = readr::read_csv("D:/LifeInUCSB/Study/GEOG176A/week2/assignment5/geog-176A-labs/data/uscities.csv")
sf_uscities = st_as_sf(uscities, coords = c("lng", "lat"), crs = 4326) %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii"))

eqdc_uscities = st_transform(sf_uscities, eqdc)
```

# Question 2

#2.1 - Distance to USA Border (coastline or national) (km)
```{r,message=FALSE, warning=FALSE}
usmapinunion = st_union(mapdata) %>%
  st_cast("MULTILINESTRING")

usmapincombined = st_combine(mapdata) %>%
  st_cast("MULTILINESTRING")

eqdc_uscities = eqdc_uscities %>% 
  mutate(distance_border = st_distance(eqdc_uscities, usmapinunion),
         distance_border = units::set_units(distance_border, "km"),
         distance_border = units::drop_units(distance_border))

eqdc_uscities %>% 
  slice_max(distance_border, n = 5) %>%
  select(city, state_name, distance_border) %>% 
  as.data.frame %>% 
  select(city, state_name, distance_border)->
  borderdis_table
knitr::kable(borderdis_table,caption = "Cities Farthest from USA Border",
               col.names = c("city","state", "distance"))
```

#2.2 - Distance to States (km)
```{r,message=FALSE, warning=FALSE}
usmap_union = st_union(mapdata) %>%
  st_cast("MULTILINESTRING")

eqdc_uscities = eqdc_uscities %>% 
  mutate(distance_state = st_distance(eqdc_uscities, usmapincombined),
         distance_state = units::set_units(distance_state, "km"),
         distance_state = units::drop_units(distance_state))

statedis_table= eqdc_uscities %>% 
  slice_max(distance_state, n = 5) %>%
  select(city, state_name, distance_state) %>% 
  as.data.frame %>% 
  select(city, state_name, distance_state)

knitr::kable(statedis_table,caption = "Cities Farthest from State Boarder",
               col.names = c("city","state", "distance"))
```

#2.3 - Distance to Mexico (km)
```{r,message=FALSE, warning=FALSE}
usmap_union = st_union(mapdata) %>%
  st_cast("MULTILINESTRING")


eqdc_uscities = eqdc_uscities %>% 
  mutate(distance_mexico = st_distance(eqdc_uscities, filter(countries_sf,admin == "Mexico")),
         distance_mexico = units::set_units(distance_mexico, "km"),
         distance_mexico = units::drop_units(distance_mexico))
eqdc_uscities %>% 
  slice_max(distance_mexico, n = 5) %>%
  select(city, state_name, distance_mexico) %>% 
  as.data.frame %>% 
  select(city, state_name, distance_mexico)->
  mexicodis_table
knitr::kable(mexicodis_table,caption = "Cities Farthest from Mexico Border",
               col.names = c("city","state", "distance"))
```

#2.4 - Distance to Canada (km)
```{r,message=FALSE, warning=FALSE}
usmap_union = st_union(mapdata) %>%
  st_cast("MULTILINESTRING")


eqdc_uscities = eqdc_uscities %>% 
  mutate(distance_canada = st_distance(eqdc_uscities, filter(countries_sf,admin == "Canada")),
         distance_canada = units::set_units(distance_canada, "km"),
         distance_canada = units::drop_units(distance_canada))

eqdc_uscities %>% 
  slice_max(distance_canada, n = 5) %>%
  select(city, state_name, distance_canada) %>% 
  as.data.frame %>% 
  select(city, state_name, distance_canada)->
  canadadis_table
knitr::kable(canadadis_table,caption = "Cities Farthest from Can Border",
               col.names = c("city","state", "distance"))
```

#Question 3

#3.1 Data

```{r,message=FALSE, warning=FALSE}
max_popu_cities = eqdc_uscities %>% 
  slice_max(population, n = 10) 

ggplot() +
  geom_sf(data = mapdata,lty = "dashed", size = 0.3) +
  geom_sf(data = countries_sf) +
  geom_sf(data = usmapincombined,size = 0.3) +
  geom_sf(data = max_popu_cities, size = 1) +
    ggrepel::geom_label_repel(
    data = max_popu_cities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2.5)
```

#3.2 City Distance from the Border
```{r,message=FALSE, warning=FALSE}
ggplot()+
  geom_sf(data = usmapinunion)+
  geom_sf(data = borderdis_table, size= 2, color = "red")+
  geom_sf(data = eqdc_uscities, aes(col = distance_border), lty = 3, alpha = .5) + 
  scale_fill_gradient(low = 'blue', high = "red",name = "Distance(km)")+
    ggrepel::geom_label_repel(
    data = borderdis_table,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
  labs(title = "City Distance from the Border of the United States",
       subtitle = "Five Cities Farthest from USA Border",
       x = "",
       y = "",
       col = "Distance from USA Border (km)") +
  theme_dark()
```

#3.3 City Distance from Nearest State
```{r,message=FALSE, warning=FALSE}
ggplot() +
  geom_sf(data = usmapinunion) +
  geom_sf(data = eqdc_uscities, aes(col = distance_state), lty = 3, alpha = .5) +
  scale_fill_gradient(low = 'green', high = "yellow",name = "Distance(km)")+
  geom_sf(data = statedis_table) +
  ggrepel::geom_label_repel(
    data = statedis_table,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 4) +
  labs(title = "City Distance from the Border of State",
       subtitle = "Five Cities Farthest from State Border",
       x = "",
       y = "",
       col = "Distance from State Border (km)") +
  theme_dark() 
```

#3.4 Equidistance boundary from Mexico and Canada
```{r,message=FALSE, warning=FALSE}
eqdc_uscities = eqdc_uscities %>% 
  mutate(distance_caminusmx = abs(distance_canada - distance_mexico),
         distance_caminusmx = units::set_units(distance_caminusmx, "km"),
         distance_caminusmx = units::drop_units(distance_caminusmx))

popucities = eqdc_uscities %>%
  filter(distance_caminusmx < 100 ) %>%
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = usmapinunion) +
  geom_sf(data = popucities)+
  geom_sf(data = eqdc_uscities, aes(col = distance_caminusmx), size = .1) +
  scale_fill_gradient(low = 'orange', high = "purple",name = "Distance(km)")+
  gghighlight::gghighlight(as.numeric(distance_caminusmx) < 100)+
  ggrepel::geom_label_repel(
    data = popucities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 4) +
  labs(title = "Cities' Distance from Nearest State Border",
       subtitle = "Five Furthest Cities from Nearest State Border",
       x = "",
       y = "",
       col = "Distance from State Border (km)") +
  theme_dark() 
```
